Psychophysics of musical elements in the discrete-time representation of sound

### 2013-09-28 article/article.tex
atualizado artigo, voltando aa traducao

The sinusoidal components are usually the main responsible for the timbre quality. If they are not presented in harmonic proportions (small number relations), the sound is perceived as noisy or dissonant, in opposite of a sonority with a unequivocally established fundamental frequency when they are present. Furthermore, the notion of absolute pitch in a complex sonority is based relies on the similarity of the spectrum to the harmonic series~\cite{Roederer} (**nao entendi o q vcs querem dizer**).

In the case of a fixed length period and waveform, the spectrum is perfectly harmonic and static. Each waveform is compound of specific proportions of harmonic components and the greater the curvature of the part, the greater the contribution of the part to the energy on the high harmonics. This aspect can be seen on real sounds. The wave labeled as ``sampled real sound'' in Figure~\ref{fig:formasDeOnda} has a period of $\Lambda_f=114$ samples, extracted from a relatively well behaved real sound. The oboe wave was sampled for A4 also in $44.1kHz$. The chose period for sampling was a relatively short one, with $98$ samples, and corresponds to the frequency $\frac{44100}{98}=450Hz$. It can be noticed from the curvatures, the oboe's rich spectrum in high frequencies and the lower spectrum of the real sound.

>>>>

The sinusoidal components are usually the main responsibles for timbre qualities. If they are not presented in harmonic proportions (small number relations), the sound is perceived as noisy or dissonant, in opposition to sonorities with an unequivocally established fundamental. Furthermore, the notion of absolute pitch relies on the similarity of the spectrum to the harmonic series~\cite{Roederer}.

In the case of a fixed length period and waveform, the spectrum is perfectly harmonic and static and  each waveform is compound of specific proportions of harmonic components. High  curvatures are sign of high harmonics in the wave. Figure~\ref{fig:formasDeOnda} depicts a wave, labeled as ``sampled real sound'', with a period of $\Lambda_f=114$ samples, extracted from a relatively well behaved recorded sound. The oboe wave was sampled for A4 also in $44.1kHz$. The chose period for sampling was a relatively short one, with $98$ samples, and corresponds to the frequency $\frac{44100}{98}=450Hz$. It can be noticed from the curvatures, the oboe's rich spectrum in high frequencies and the lower spectrum of the real sound.
